---
title: "Lab 1: Intro to R"
output: 
  html_document: 
    highlight: pygments
    theme: spacelab
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

* * *

## Part I: Introduction to R

(optional) If you're new to R, please work through the lab in section 2.3 of your
there is no need to include that code in your lab report.

* * *

## Part II: Boston Housing Data

(based on exercise 10 on p. 56 of the book)

To begin, load in the `Boston` data set. The `Boston` data set is part of the
`MASS` library in R.

```{r}
library(MASS)
```

Now the data set is contained in the object `Boston`. Read about the data set.

```{r}
?Boston
```

1. How many rows are in this data set? How many columns? What do the rows and columns
represent?

2. Make some (2-3) pairwise scatterplots of the predictors (columns) in this data set.
Describe your findings.

3. Are any of the predictors associated with per capita crime rate? If so, explain
the relationship.

4. Are there any suburbs of Boston that appear to have particularly high crime rates?
Tax rate? Pupil-teacher ratios? Comment on the range of each predictor.

5. How many of the suburbs in this data set bound the Charles river?

6. What is the median pupil-teacher ratio among the towns in this data set?

7. If you want to build a model to predict the average value of a home based on
the other variables, what is your output/response? What is your input?



```{r, echo = FALSE, eval = FALSE}
## Part II: Simulation

R is terrific at simulating data from a model of your specifications. Say I want
a model that expresses the idea that we have two sub-populations of whales that
differ in their length and their coloration (measured by mean
grayscale value). I can create my own data set for a hypothetical group of whales
using the following. Please run each line one by one and consider the output of
each (you can view each new object by clicking on it in the Environment pane).
n_whales <- 40
group <- rep(c(1, 2), each = n_whales/2)
length_1 <- rnorm(n = n_whales/2, mean = 4.2, sd = .4)
length_2 <- rnorm(n = n_whales/2, mean = 3.6, sd = .3)
len <- c(length1, length2)
color_1 <- rnorm(n = n_whales/2, mean = 4.2, sd = .4)

x <- seq(0, 100, )
```



