---
title: "Hierarchical Clustering"
output:
  ioslides_presentation:
    incremental: true
---

```{r echo = FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE)
```

## Algorithm 10.2 {.build}

1. Begin with $n$ observations and a dissimilarity matrix. Treat each observation as its own cluster.
2. For $i = n, n - 1, \ldots, 2$:
    - Examine all pairwise inter-cluster dissimilarities and identify the pair that are most similar. Fuse these clusters at the dendrogram height corresponding to their dissimilarity.
    - Compute pairwise dissimilarities among remaining 1 - i clusters.
    

## Linkage types {.build}

**Complete**: Maximal intercluster dissimilarity.

**Single**: Minimal intercluster dissimilarity.

**Average**: Mean intercluster dissimiliarity.

**Centroid**: Dissimilarity between cluster centroids.


#

## Example: Cancer Genomics {.build}

<img src = "http://mkweb.bcgsc.ca/dev/circos//img/masthead/circos-embl-cancer-genomics-2013.png" width = 750>

How do you characterize the gene expression in cancerous cells?

`NCI60`: Cancer cell mircoarray data with 6,830 expression measurements on 64 cancer cell lines.


## {.build}
```{r}
library(ISLR)
NCI <- NCI60$data
dim(NCI)
NCI[1:3, 1:5]
```

## {.build}

```{r}
NCI_scaled <- scale(NCI)
NCI_dist <- dist(NCI_scaled)
dissmatrix <- as.matrix(NCI_dist)
dissmatrix[1:5, 1:5]
```

This distance matrix can be visualized as a *heatmap*.


## {.build}

```{r fig.align='center', fig.width = 6, fig.height = 6}
heatmap(dissmatrix, Rowv = NA, Colv = NA)
```


## {.build}

```{r fig.align='center', fig.width = 6, fig.height = 6}
heatmap(dissmatrix, Colv = NA)
```


## Dendrogram {.build}

```{r, echo = FALSE, fig.width = 8, fig.align='center'}
plot(hclust(NCI_dist), xlab = "", ylab = "", sub = "", main = "Complete Linkage", cex = .6)
```


## Dendrogram

```{r, echo = FALSE, fig.width = 8, fig.align='center'}
plot(hclust(NCI_dist), labels = NCI60$labs, xlab = "", ylab = "", sub = "", main = "Complete Linkage", cex = .6)
```


## Dendrogram

```{r, echo = FALSE, fig.width = 8, fig.align='center'}
plot(hclust(NCI_dist, method = "single"), labels = NCI60$labs, xlab = "", ylab = "", sub = "", main = "Single Linkage", cex = .6)
```


## Cutting the Dendro {.build}

```{r}
hc <- hclust(NCI_dist)
hc_cut <- cutree(hc, k = 4)
length(hc_cut)
head(hc_cut)
```


## Cutting the Dendro {.build}

```{r}
table(hc_cut, NCI60$labs)
```

- Leukemia cell lines in cluster 3.
- Breast cancer cell lines spread across 1, 2, and 4.


## Dendrogram

```{r, echo = FALSE, fig.width = 8, fig.align='center'}
plot(hclust(NCI_dist), labels = NCI60$labs, xlab = "", ylab = "", sub = "", main = "Complete Linkage", cex = .6)
abline(h = 139, col = "tomato")
```



