---
layout: page
title: Problem Set 2 Key
permalink: /problem-sets/
published: FALSE
output: 
  html_document: 
    highlight: pygments
    theme: spacelab
---

#### Exercise 1

```{r}
data(quakes)
plot(stations ~ jitter(mag), data = quakes, xlab = "magnitude")
```

There is a fairly strong positive relationship between magnitude and stations. The shape of the relationship is linear with potentially slight curvature.

#### Exercise 2
If there were no relationship I would expect the slope to be 0, and the intercept to be the mean value of stations.

#### Exercise 3
```{r}
m1 <- lm(stations ~ mag, data = quakes)
plot(stations ~ jitter(mag), data = quakes, xlab = "magnitude")
abline(m1)
summary(m1)
```
For this line, the slope indicates the number of additional stations that would
be expected to detect an earthquake that is one unit greater in magnitude. In this case, the intercept is the number of stations that would detect a magnitude 0 event. However, since these events cannot be detected, and furthermore the data doesn't go below magnitude 0 events, this does not correspond to a real value, so the fact that it's negative shouldn't bother us.

#### Exercise 4
To manually calculate the slope, I use the formula from the book:
```{r}
s <- quakes$stations
sb <- mean(s)
m <- quakes$mag
mb <- mean(m)
b1 <- sum((s - sb)*(m - mb))/ sum((m - mb)^2)
b1
```
This is the exact same value as given by `lm()`, so it checks out.

#### Exercise 5
Using Eq. 3.9 from the book, our interval is given by

$$ \hat \beta_1 \pm 2 \cdot SE(\hat \beta_1), $$

where our standard error is given by

$$ SE(\hat \beta_1) = \dfrac{\sigma} {\sqrt{\sum (x_i - \bar x)^2}} $$

but summary() gives the standard error of our slope to be 0.903, so let's not calculate it over again.

```{r}
SEb1 <- summary(m1)$coefficients[2,2]
b1_upper <- b1 + 2 * SEb1
b1_lower <- b1 - 2 * SEb1
b1_upper
b1_lower
confint(m1)
```

So our results are close, but strangely enough not exactly the same. I expect they would be closer if I used the exact t-value instead of 2.

#### Exercise 6
Plugging in mag = 7 to our model:

```{r}
 m1$coefficients[1] + 7 * m1$coefficients[2]
```

So we would expect about 143 detections for a magnitude 7 earthquake.

#### Exercise 7
1. We described the data by characterizing the relationship.
2. Inference, since we're considering an unknown true parameter.
3. Even though we've fit a model, the slope and intercept can be thought of
as geometric descriptions of the data.
4. Same as 3.
5. Inference on a true slope parameter.
6. This is an attempt at prediction, which to be accurate, needs to draw lessons from inference in thinking about capturing true parameters and not overfitting our data.

#### Exercise 9
```{r}
x <- quakes$mag
```
For this section, I simply reused the magnitude data from the actual dataset.

#### Exercise 10
```{r}
f_hat <- function(x)
{ m1$coefficients[1] + x * m1$coefficients[2]}
y_hat <- f_hat(x)
```

#### Exercise 11
```{r}
rss <- sum(m1$residuals^2)
sigma2 <- rss/(length(m1$residuals) - 2)
y <- y_hat + rnorm(length(x), sd = sqrt(sigma2)) 
```


#### Exercise 12
```{r}
plot(x, y)
```

This plot looks pretty close to the real one! The real data has a bit more of an upward curve which the linear model can't capture, but the spread of the data is the same.
