---
title: "Clustering: choosing K"
output:
  ioslides_presentation:
    incremental: true
---

```{r echo = FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE)
```

## {.build}

```{r}
poverty <- read.delim("https://raw.githubusercontent.com/andrewpbray/math-243/gh-pages/assets/data/poverty.txt", header = TRUE)
km1 <- kmeans(poverty[ ,-1], 4, nstart = 20)
library(dplyr)
kd <- arrange(data.frame(poverty$State, "k" = km1$cluster), k)
head(kd)
tail(kd)
```


## PC representation

```{r eval = FALSE}
pca1 <- prcomp(poverty[, -1])
d <- data.frame(PC1 = pca1$x[, 1],
                PC2 = pca1$x[, 2],
                cluster = as.factor(km1$cluster),
                state = poverty$State)
ggplot(d, aes(x = PC1, y = PC2, col = cluster)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```


## K = 4

```{r, echo = FALSE, fig.align='center', fig.width=8, fig.height=5}
set.seed(40)
km4 <- kmeans(poverty[ ,-1], 4, nstart = 20)
pca1 <- prcomp(poverty[, -1])
d <- data.frame(PC1 = pca1$x[, 1],
                PC2 = pca1$x[, 2],
                cluster = as.factor(km4$cluster),
                state = poverty$State)
library(ggplot2)

ggplot(d, aes(x = PC1, y = PC2, col = cluster)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```


## K = 3

```{r, echo = FALSE, fig.align='center', fig.width=8, fig.height=5}
set.seed(40)
km3 <- kmeans(poverty[ ,-1], 3, nstart = 20)
d <- data.frame(PC1 = pca1$x[, 1],
                PC2 = pca1$x[, 2],
                cluster = as.factor(km3$cluster),
                state = poverty$State)
library(ggplot2)

ggplot(d, aes(x = PC1, y = PC2, col = cluster)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```


## K = 2

```{r, echo = FALSE, fig.align='center', fig.width=8, fig.height=5}
set.seed(40)
km2 <- kmeans(poverty[ ,-1], 2, nstart = 20)
d <- data.frame(PC1 = pca1$x[, 1],
                PC2 = pca1$x[, 2],
                cluster = as.factor(km2$cluster),
                state = poverty$State)
library(ggplot2)

ggplot(d, aes(x = PC1, y = PC2, col = cluster)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```


## K = 1

```{r, echo = FALSE, fig.align='center', fig.width=8, fig.height=5}
set.seed(40)
km1 <- kmeans(poverty[ ,-1], 1, nstart = 20)
d <- data.frame(PC1 = pca1$x[, 1],
                PC2 = pca1$x[, 2],
                cluster = as.factor(km1$cluster),
                state = poverty$State)
library(ggplot2)

ggplot(d, aes(x = PC1, y = PC2, col = cluster)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```


## Variation with K = 1 {.build}

```{r}
names(km1)
km1$withinss
km1$tot.withinss
```


## K = 2

```{r, echo = FALSE, fig.align='center', fig.width=8, fig.height=5}
set.seed(40)
km2 <- kmeans(poverty[ ,-1], 2, nstart = 20)
d <- data.frame(PC1 = pca1$x[, 1],
                PC2 = pca1$x[, 2],
                cluster = as.factor(km2$cluster),
                state = poverty$State)
library(ggplot2)

ggplot(d, aes(x = PC1, y = PC2, col = cluster)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```


## Variation with K = 2 {.build}

```{r}
km2$withinss
km2$tot.withinss
km2$totss
```


## TWSS and K

```{r echo = FALSE}
km5 <- kmeans(poverty[ ,-1], 5, nstart = 20)
km6 <- kmeans(poverty[ ,-1], 6, nstart = 20)
km7 <- kmeans(poverty[ ,-1], 7, nstart = 20)
km8 <- kmeans(poverty[ ,-1], 8, nstart = 20)
df <- data.frame("K" = 1:8,
                 "SS" = c(km1$tot.withinss,
                          km2$tot.withinss,
                          km3$tot.withinss,
                          km4$tot.withinss,
                          km5$tot.withinss,
                          km6$tot.withinss,
                          km7$tot.withinss,
                          km8$tot.withinss))
ggplot(df, aes(x = K, y = SS)) +
  geom_point() +
  geom_line()
```


## Selecting K {.build}

- Use domain area knowledge.
- Look for "elbow" in a scree plot.
- Formalize "elbow" with Gap statistic (Tibshirani, 2001).

The number of clusters is often ambiguous, which shouldn't be surprising in an unsupervised setting.

Choice of K is choosing where on the spectrum between complete aggregation (K = 1) and no aggregation (K = n).




